---
title: "IDS investigation worksheet"
author: "by Team-Name: User1, User2, User3, User4 & User5."
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your 
group project.

5 Steps 

- stating and refining question. k

  - Describe - 
  - Explore - analyse for relationships between variables, crime and community variables. 
  - Infer - Analyse data from a group, race, age, income, parents
  - Predict - Predictions on an individual or a population. Must involve the goal variable, Violent crimes per population. A look at the determining factors of Violent Crimes Per Pop
  - Casual - 
  - Mechanistic - how does one variable impact another
  
  - confounding variables?
  
- exploratory data analysis

  - graph against violent crimes per population 
  - frequency distribution
  
  checklist:
  
  - validate
  - top,  bottom, middle rows
  - plot
  - 

- model building

- interpreting results

- communicating results

Data Type, attributes, kinds of questions i can answer.

Objective Statement/ Problem Statement

cleaning 
- NA
- column name
- importing
- remove 0 and 1 from dataset and replace with... 

back up data source


Exploratory Data Analysis:

-What type of variation occurs within variables?

  - what is most common? why?
  - what values are rare? why? does this match expectations?
  - unusual patterns? Explain
  - How are obsrvations within subgroup similar
  - how are observations in different clusters diffeerent from each other
  - explain or descrie clusters
  
See a Pattern? Ask yourself:

- is it coincidence?
- how can you describe the relationship?
- how strong is the relationship
- what other variables may affect the relationship
- does the relationship change if you look at subgroups of the data

-What type of covariation occurs between my variables?




summary statistics
- Mean 
- Standard deviation

visualisations

map

select algorithm/method

evaluate model

The Question.

What is Crime?

- 

What is community?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
# Add any other libraries here



```


```{r load-data, include = TRUE}
# load your data 

data <- read.csv("data/crime/communitiesUnnormalized.txt", header = FALSE, na.string = "?")
  
names_lines <- read_lines("data/crime/names.names")
  
  
attribute_lines <- grep("^@attribute", names_lines, value = TRUE)
attribute_lines <- gsub("@attribute ", "", attribute_lines)
attribute_lines <- gsub(" .*", "", attribute_lines)

colnames(data) <- c(attribute_lines)

column_names <- colnames(data)

```

```{r main-program}

# Poverty vs crime

data <- data %>%
  mutate(
    incomeLevel = cut(medIncome, breaks = 3, labels = c("Low", "Middle", "High"), include.lowest = TRUE)
    )

ggplot(data, aes(x = incomeLevel, y = violentPerPop)) +
  geom_boxplot(fill = "orange", color = "black") +
  labs(
    title = "Relation between income level and crimes per population",
    subtitle = "",
    x = "Income Level",
    y = "Violent Crimes"
  ) +
  theme_bw()

ggplot(data, aes(x = pctPoverty, y = violentPerPop)) +
  geom_point(color = "orange") +
  geom_smooth(method = lm, se = FALSE) +
  theme_minimal()

# Family vs crime

family_statistics <- data %>%
  select(medFamIncome, pct2Par, `pctWorkMom-6`, violentPerPop) %>%
  drop_na()

family_statistics <- family_statistics %>%
  rename(
    "Median Family Income" = medFamIncome,
    "Percentage of families with two parents" = pct2Par,
    "Percentage of working moms with young kids" = `pctWorkMom-6`
  )

# Geography vs crime

# Install and load required packages
if (!requireNamespace("leaflet", quietly = TRUE)) {
  install.packages("leaflet")
}
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
# Load required libraries
library(tidyverse)
library(leaflet)
 
data <- read.csv("data/crime/communities.data", header = FALSE, stringsAsFactors = FALSE, na.strings = "?")
# Load attribute names
data_names <- read_lines("data/crime/communities.names")
attribute_lines <- grep("^@attribute", data_names, value = TRUE)
attribute_lines <- gsub("@attribute ", "", attribute_lines)
attribute_lines <- gsub(" .*", "", attribute_lines)
colnames(data) <- attribute_lines
# Add latitude and longitude columns manually
state_names <- data.frame(
  state = c(1, 2, 4, 5, 6, 8, 9, 10, 11, 12, 13, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 44, 45, 46, 47, 48, 49, 50, 51, 53, 54, 55, 56),
  state_name = c("ALABAMA", "ALASKA", "ARIZONA", "ARKANSAS", "CALIFORNIA", "COLORADO", "CONNECTICUT", "DELAWARE", "DISTRICT OF COLUMBIA", "FLORIDA", "GEORGIA", "HAWAII", "IDAHO", "ILLINOIS", "INDIANA", "IOWA", "KANSAS", "KENTUCKY", "LOUISIANA", "MAINE", "MARYLAND", "MASSACHUSETTS", "MICHIGAN", "MINNESOTA", "MISSISSIPPI", "MISSOURI", "MONTANA", "NEBRASKA", "NEVADA", "NEW HAMPSHIRE", "NEW JERSEY", "NEW MEXICO", "NEW YORK", "NORTH CAROLINA", "NORTH DAKOTA", "OHIO", "OKLAHOMA", "OREGON", "PENNSYLVANIA", "RHODE ISLAND", "SOUTH CAROLINA", "SOUTH DAKOTA", "TENNESSEE", "TEXAS", "UTAH", "VERMONT", "VIRGINIA", "WASHINGTON", "WEST VIRGINIA", "WISCONSIN", "WYOMING"),
  latitude = c(32.806671, 61.370716, 33.729759, 34.969704, 36.116203, 39.059811, 41.597782, 39.318523, 38.897438, 27.766279, 33.040619, 21.094318, 44.240459, 40.349457, 39.849426, 42.011539, 38.5266, 37.66814, 31.169546, 44.693947, 39.063946, 42.230171, 43.326618, 45.694454, 32.741646, 38.456085, 46.921925, 41.12537, 38.313515, 43.452492, 40.298904, 34.840515, 42.165726, 35.630066, 47.528912, 40.388783, 35.565342, 44.572021, 40.590752, 41.680893, 33.856892, 44.299782, 35.747845, 31.054487, 39.32098, 44.045876, 37.769337, 47.400902, 38.491226, 44.268543, 42.755966),
  longitude = c(-86.79113, -152.404419, -111.431221, -92.373123, -119.681564, -105.311104, -72.755371, -75.507141, -77.026817, -81.686783, -83.643074, -157.498337, -114.478828, -88.986137, -86.258278, -93.210526, -96.726486, -84.670067, -91.867805, -69.381927, -76.802101, -71.530106, -84.536095, -93.900192, -89.678696, -92.288368, -110.454353, -98.268082, -117.055374, -71.563896, -74.521011, -106.248482, -74.948051, -79.806419, -99.784012, -82.764915, -96.928917, -122.070938, -77.209755, -71.51178, -80.945007, -99.438828, -86.692345, -97.563461, -111.950684, -72.710686, -78.169968, -121.490494, -80.954456, -89.616508, -107.30249)
)
# Join state information with data
data <- data %>%
  left_join(state_names, by = "state") %>%
  select(state_name, communityname, ViolentCrimesPerPop, latitude, longitude, everything())
# Filter and prepare the data
map_data <- data %>%
  filter(!is.na(latitude) & !is.na(longitude) & !is.na(ViolentCrimesPerPop)) %>%
  select(communityname, ViolentCrimesPerPop, latitude, longitude)
# Create a leaflet map
data_map <- leaflet(map_data) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
lng = ~longitude, lat = ~latitude,
    radius = ~pmax(ViolentCrimesPerPop * 10, 3),
    color = "red",
    stroke = FALSE,
    fillOpacity = 0.5,
    popup = ~paste("Community: ", communityname, "<br>",
                   "Violent Crimes per Pop: ", ViolentCrimesPerPop)
  )
 
data_map



# Install and load required packages
if (!requireNamespace("leaflet", quietly = TRUE)) {
  install.packages("leaflet")
}
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
# Load required libraries
library(tidyverse)
library(leaflet)
 
data <- read.csv("data/crime/communities.data", header = FALSE, stringsAsFactors = FALSE, na.strings = "?")
# Load attribute names
data_names <- read_lines("data/crime/communities.names")
attribute_lines <- grep("^@attribute", data_names, value = TRUE)
attribute_lines <- gsub("@attribute ", "", attribute_lines)
attribute_lines <- gsub(" .*", "", attribute_lines)
colnames(data) <- attribute_lines
# Add latitude and longitude columns manually
state_names <- data.frame(
  state = c(1, 2, 4, 5, 6, 8, 9, 10, 11, 12, 13, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 44, 45, 46, 47, 48, 49, 50, 51, 53, 54, 55, 56),
  state_name = c("ALABAMA", "ALASKA", "ARIZONA", "ARKANSAS", "CALIFORNIA", "COLORADO", "CONNECTICUT", "DELAWARE", "DISTRICT OF COLUMBIA", "FLORIDA", "GEORGIA", "HAWAII", "IDAHO", "ILLINOIS", "INDIANA", "IOWA", "KANSAS", "KENTUCKY", "LOUISIANA", "MAINE", "MARYLAND", "MASSACHUSETTS", "MICHIGAN", "MINNESOTA", "MISSISSIPPI", "MISSOURI", "MONTANA", "NEBRASKA", "NEVADA", "NEW HAMPSHIRE", "NEW JERSEY", "NEW MEXICO", "NEW YORK", "NORTH CAROLINA", "NORTH DAKOTA", "OHIO", "OKLAHOMA", "OREGON", "PENNSYLVANIA", "RHODE ISLAND", "SOUTH CAROLINA", "SOUTH DAKOTA", "TENNESSEE", "TEXAS", "UTAH", "VERMONT", "VIRGINIA", "WASHINGTON", "WEST VIRGINIA", "WISCONSIN", "WYOMING"),
  latitude = c(32.806671, 61.370716, 33.729759, 34.969704, 36.116203, 39.059811, 41.597782, 39.318523, 38.897438, 27.766279, 33.040619, 21.094318, 44.240459, 40.349457, 39.849426, 42.011539, 38.5266, 37.66814, 31.169546, 44.693947, 39.063946, 42.230171, 43.326618, 45.694454, 32.741646, 38.456085, 46.921925, 41.12537, 38.313515, 43.452492, 40.298904, 34.840515, 42.165726, 35.630066, 47.528912, 40.388783, 35.565342, 44.572021, 40.590752, 41.680893, 33.856892, 44.299782, 35.747845, 31.054487, 39.32098, 44.045876, 37.769337, 47.400902, 38.491226, 44.268543, 42.755966),
  longitude = c(-86.79113, -152.404419, -111.431221, -92.373123, -119.681564, -105.311104, -72.755371, -75.507141, -77.026817, -81.686783, -83.643074, -157.498337, -114.478828, -88.986137, -86.258278, -93.210526, -96.726486, -84.670067, -91.867805, -69.381927, -76.802101, -71.530106, -84.536095, -93.900192, -89.678696, -92.288368, -110.454353, -98.268082, -117.055374, -71.563896, -74.521011, -106.248482, -74.948051, -79.806419, -99.784012, -82.764915, -96.928917, -122.070938, -77.209755, -71.51178, -80.945007, -99.438828, -86.692345, -97.563461, -111.950684, -72.710686, -78.169968, -121.490494, -80.954456, -89.616508, -107.30249)
)
# Join state information with data
data <- data %>%
  left_join(state_names, by = "state") %>%
  select(state_name, communityname, PolicPerPop, latitude, longitude, everything())
# Filter and prepare the data
map_data <- data %>%
  filter(!is.na(latitude) & !is.na(longitude) & !is.na(PolicPerPop)) %>%
  select(communityname, PolicPerPop, latitude, longitude)
# Create a leaflet map
data_map <- leaflet(map_data) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
lng = ~longitude, lat = ~latitude,
    radius = ~pmax(PolicPerPop * 10, 3),
    color = "blue",
    stroke = FALSE,
    fillOpacity = 0.5,
    popup = ~paste("Community: ", communityname, "<br>",
                   "Police per pop: ", PolicPerPop)
  )
 
data_map








 
if (!requireNamespace("leaflet", quietly = TRUE)) {
  install.packages("leaflet")
}
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
# Load required libraries
library(tidyverse)
library(leaflet)
# Load the data
data <- read.csv("data/crime/communities.data", header = FALSE, stringsAsFactors = FALSE, na.strings = "?")
# Load attribute names
data_names <- read_lines("data/crime/communities.names")
attribute_lines <- grep("^@attribute", data_names, value = TRUE)
attribute_lines <- gsub("@attribute ", "", attribute_lines)
attribute_lines <- gsub(" .*", "", attribute_lines)
colnames(data) <- attribute_lines
# Add latitude and longitude columns manually
state_names <- data.frame(
  state = c(1, 2, 4, 5, 6, 8, 9, 10, 11, 12, 13, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 44, 45, 46, 47, 48, 49, 50, 51, 53, 54, 55, 56),
  state_name = c("ALABAMA", "ALASKA", "ARIZONA", "ARKANSAS", "CALIFORNIA", "COLORADO", "CONNECTICUT", "DELAWARE", "DISTRICT OF COLUMBIA", "FLORIDA", "GEORGIA", "HAWAII", "IDAHO", "ILLINOIS", "INDIANA", "IOWA", "KANSAS", "KENTUCKY", "LOUISIANA", "MAINE", "MARYLAND", "MASSACHUSETTS", "MICHIGAN", "MINNESOTA", "MISSISSIPPI", "MISSOURI", "MONTANA", "NEBRASKA", "NEVADA", "NEW HAMPSHIRE", "NEW JERSEY", "NEW MEXICO", "NEW YORK", "NORTH CAROLINA", "NORTH DAKOTA", "OHIO", "OKLAHOMA", "OREGON", "PENNSYLVANIA", "RHODE ISLAND", "SOUTH CAROLINA", "SOUTH DAKOTA", "TENNESSEE", "TEXAS", "UTAH", "VERMONT", "VIRGINIA", "WASHINGTON", "WEST VIRGINIA", "WISCONSIN", "WYOMING"),
  latitude = c(32.806671, 61.370716, 33.729759, 34.969704, 36.116203, 39.059811, 41.597782, 39.318523, 38.897438, 27.766279, 33.040619, 21.094318, 44.240459, 40.349457, 39.849426, 42.011539, 38.5266, 37.66814, 31.169546, 44.693947, 39.063946, 42.230171, 43.326618, 45.694454, 32.741646, 38.456085, 46.921925, 41.12537, 38.313515, 43.452492, 40.298904, 34.840515, 42.165726, 35.630066, 47.528912, 40.388783, 35.565342, 44.572021, 40.590752, 41.680893, 33.856892, 44.299782, 35.747845, 31.054487, 39.32098, 44.045876, 37.769337, 47.400902, 38.491226, 44.268543, 42.755966),
  longitude = c(-86.79113, -152.404419, -111.431221, -92.373123, -119.681564, -105.311104, -72.755371, -75.507141, -77.026817, -81.686783, -83.643074, -157.498337, -114.478828, -88.986137, -86.258278, -93.210526, -96.726486, -84.670067, -91.867805, -69.381927, -76.802101, -71.530106, -84.536095, -93.900192, -89.678696, -92.288368, -110.454353, -98.268082, -117.055374, -71.563896, -74.521011, -106.248482, -74.948051, -79.806419, -99.784012, -82.764915, -96.928917, -122.070938, -77.209755, -71.51178, -80.945007, -99.438828, -86.692345, -97.563461, -111.950684, -72.710686, -78.169968, -121.490494, -80.954456, -89.616508, -107.30249)
)
# Join state information with data
data <- data %>%
  left_join(state_names, by = "state")
# Filter and prepare the data
map_data <- data %>%
  filter(!is.na(latitude) & !is.na(longitude) & !is.na(PolicPerPop) & !is.na(ViolentCrimesPerPop)) %>%
  select(state, communityname, ViolentCrimesPerPop, PolicPerPop, latitude, longitude)
# Create a leaflet map
data_map <- leaflet(map_data) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%  # Use a different tile provider for better visualization
  addCircleMarkers(
    lng = ~longitude, lat = ~latitude,
    radius = ~pmax(ViolentCrimesPerPop * 10, 3),  # Ensure minimum radius for visibility
    color = "red",
    stroke = FALSE,
    fillOpacity = 0.5,
    popup = ~paste("Community: ", communityname, "<br>",
                   "Violent Crimes per Pop: ", ViolentCrimesPerPop)
  ) %>%
  addCircleMarkers(
    lng = ~longitude, lat = ~latitude,
    radius = ~pmax(PolicPerPop * 10, 3),  # Ensure minimum radius for visibility
    color = "blue",
    stroke = FALSE,
    fillOpacity = 0.5,
    popup = ~paste("Community: ", communityname, "<br>",
                   "Police Per Pop: ", PolicPerPop)
  )
# Print the map
data_map





```

```{r}
# Geospatial data available at the geojson format
tmp_geojson <- tempfile(fileext = ".geojson")

# Download the GeoJSON file (note: use raw URL for correct download)
download.file(
  "https://raw.githubusercontent.com/PublicaMundi/MappingAPI/master/data/geojson/us-states.json",
  tmp_geojson
)

# Load necessary package
library(sf)

# Read the GeoJSON file
my_sf <- read_sf(tmp_geojson)

library(ggplot2)
ggplot(my_sf) +
  geom_sf(fill = "white", color = "black", linewidth = 0.3) +
  theme_void()


```

```{r}
# Geospatial data available at the geojson format
tmp_geojson <- tempfile(fileext = ".geojson")

# Download the GeoJSON file (note: use raw URL for correct download)
download.file(
  "https://raw.githubusercontent.com/PublicaMundi/MappingAPI/master/data/geojson/us-states.json",
  tmp_geojson
)

# Load necessary packages
library(sf)
library(ggplot2)
library(dplyr)

# Read the GeoJSON file
my_sf <- read_sf(tmp_geojson)

# Example dataset 'violentPerPop' with state codes and names
violentPerPop <- data.frame(
  state = c(1, 2, 4, 5, 6, 8, 9, 10, 11, 12, 13, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 44, 45, 46, 47, 48, 49, 50, 51, 53, 54, 55, 56),
  state_name = c("ALABAMA", "ALASKA", "ARIZONA", "ARKANSAS", "CALIFORNIA", "COLORADO", "CONNECTICUT", "DELAWARE", "DISTRICT OF COLUMBIA", "FLORIDA", "GEORGIA", "HAWAII", "IDAHO", "ILLINOIS", "INDIANA", "IOWA", "KANSAS", "KENTUCKY", "LOUISIANA", "MAINE", "MARYLAND", "MASSACHUSETTS", "MICHIGAN", "MINNESOTA", "MISSISSIPPI", "MISSOURI", "MONTANA", "NEBRASKA", "NEVADA", "NEW HAMPSHIRE", "NEW JERSEY", "NEW MEXICO", "NEW YORK", "NORTH CAROLINA", "NORTH DAKOTA", "OHIO", "OKLAHOMA", "OREGON", "PENNSYLVANIA", "RHODE ISLAND", "SOUTH CAROLINA", "SOUTH DAKOTA", "TENNESSEE", "TEXAS", "UTAH", "VERMONT", "VIRGINIA", "WASHINGTON", "WEST VIRGINIA", "WISCONSIN", "WYOMING"),
  value = runif(51, min = 0, max = 10)  # Replace with actual data if available
)

# Join the GeoJSON spatial data with the violentPerPop dataset
choropleth_data <- my_sf %>%
  left_join(violentPerPop, by = c("name" = "state_name"))

# Create a choropleth map using ggplot2 with a red color scale
ggplot(data = choropleth_data) +
  geom_sf(aes(fill = value)) +
  scale_fill_gradient(low = "lightcoral", high = "darkred", na.value = "grey50", name = "Violence Rate") +
  theme_minimal() +
  labs(
    title = "Choropleth Map of Violence Rates by State",
    fill = "Violence Rate"
  ) +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )

```

```{r}

# Geospatial data available at the geojson format
tmp_geojson <- tempfile(fileext = ".geojson")

# Download the GeoJSON file (note: use raw URL for correct download)
download.file(
  "https://raw.githubusercontent.com/PublicaMundi/MappingAPI/master/data/geojson/us-states.json",
  tmp_geojson
)

# Load necessary packages
library(sf)
library(ggplot2)
library(dplyr)

# Read the GeoJSON file
my_sf <- read_sf(tmp_geojson)

# Remove Alaska and Hawaii from the GeoJSON data
my_sf <- my_sf %>%
  filter(!(name %in% c("ALASKA", "HAWAII")))

# Ensure state names in my_sf are uppercase to match the violentPerPop dataset
my_sf$name <- toupper(my_sf$name)


# Remove Alaska and Hawaii from the violentPerPop dataset
violentPerPop <- violentPerPop %>%
  filter(!(state_name %in% c("ALASKA", "HAWAII")))

# Join the GeoJSON spatial data with the violentPerPop dataset
choropleth_data <- my_sf %>%
  left_join(violentPerPop, by = c("name" = "state_name"))
  

# Create a choropleth map using ggplot2 with a customized style
p <- ggplot(choropleth_data) +
  geom_sf(aes(fill = value), linewidth = 0, alpha = 0.9) +
  theme_void() +
  scale_fill_gradient(
    low = "lightcoral", high = "darkred",
    na.value = "grey50",
    name = "Violence Rate",
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(12, units = "mm"),
      label.position = "bottom",
      title.position = "top",
      nrow = 1
    )
  ) +
  labs(
    title = "Choropleth Map of Violence Rates by State",
    subtitle = "Violence rates per state based on provided data",
    caption = "Data: User Dataset | Creation: Adapted by User"
  ) +
  theme(
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(
      size = 20, hjust = 0.01, color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.4, l = 2,
        unit = "cm"
      )
    ),
    plot.subtitle = element_text(
      size = 15, hjust = 0.01,
      color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.43, l = 2,
        unit = "cm"
      )
    ),
    plot.caption = element_text(
      size = 10,
      color = "#4e4d47",
      margin = margin(
        b = 0.3, r = -99, t = 0.3,
        unit = "cm"
      )
    ),
    legend.position = c(0.7, 0.09)
  )





```
```{r}

# Geospatial data available at the geojson format
tmp_geojson <- tempfile(fileext = ".geojson")

# Download the GeoJSON file (note: use raw URL for correct download)
download.file(
  "https://raw.githubusercontent.com/PublicaMundi/MappingAPI/master/data/geojson/us-states.json",
  tmp_geojson
)

# Load necessary packages
library(sf)
library(ggplot2)
library(dplyr)

# Read the GeoJSON file
my_sf <- read_sf(tmp_geojson)

# Ensure state names in my_sf are uppercase to match the violentPerPop dataset
my_sf$name <- toupper(my_sf$name)

# Join the GeoJSON spatial data with the violentPerPop dataset
choropleth_data <- my_sf %>%
  left_join(violentPerPop, by = c("name" = "state_name"))

# Create a choropleth map using ggplot2 with a customized style
p <- ggplot(choropleth_data) +
  geom_sf(aes(fill = value), linewidth = 0, alpha = 0.9) +
  theme_void() +
  scale_fill_gradient(
    low = "lightcoral", high = "darkred",
    na.value = "grey50",
    name = "Violence Rate",
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(12, units = "mm"),
      label.position = "bottom",
      title.position = "top",
      nrow = 1
    )
  ) +
  labs(
    title = "Map of the USA Showing Violent Crimes per 100k",
    subtitle = "",
    caption = "Data: violentPerPop"
  ) +
  theme(
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    panel.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(
      size = 13, hjust = 0.01, color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.4, l = 2,
        unit = "cm"
      )
    ),
    plot.subtitle = element_text(
      size = 10, hjust = 0.01,
      color = "#4e4d47",
      margin = margin(
        b = -0.1, t = 0.43, l = 2,
        unit = "cm"
      )
    ),
   
    legend.position = c(0.7, 0.09)
  )



```

