---
title: "IDS investigation worksheet"
author: "by Team-Name: User1, User2, User3, User4 & User5"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.

What is Crime?

What is comunity?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
# Add any other libraries here



```


```{r load-data}
# load your data 

data <- read.csv("data/crime/communities.data", header = FALSE, na.string = "?")
  
names_lines <- read_lines("data/crime/communities.names")
  
  
attribute_lines <- grep("^@attribute", names_lines, value = TRUE)
attribute_lines <- gsub("@attribute ", "", attribute_lines)
attribute_lines <- gsub(" .*", "", attribute_lines)
  
colnames(data) <- c(attribute_lines)

frequency_state <- data %>%
  group_by(state) %>%
  summarise(count = n())

frequency_community <- data %>%
  group_by(communityname) %>%
  summarise(count = n())


# view column names

column_names <- colnames(data)

view(column_names)

name <- list()

for (x in length(column_names)) {
  name_block <- c(name, name_block)
}

view(name_block)

#Race

race <- data %>%
  select("state", "communityname",  "population","racepctblack","racePctWhite","racePctAsian","racePctHisp") 
  view(race)


race(data)
#Age


  
age <- data %>%
  select("state", "communityname", "population", "agePct12t21", "agePct12t29", "agePct16t24", "agePct65up")
  view(age)


age(data)

# employment
employment_freq <- data %>%
  select()


employment_freq(data)
#education
  
#per capita
  
#poverty
  
#divorce etc
```

```{r load-data} 
# Main Program




```